### 2012-03-19 main.tex
changed from metric to crtieria

\documentclass{llncs}

\usepackage{url}

\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{booktabs}

\usepackage{color}

\newcommand{\todo}[2]{\textbf{\textcolor{red}{(TODO [#1]: #2)}}}

% metrics = quantitative indicator of sth

\title{Quality Criteria for SKOS Vocabularies}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2}}
\institute{
	University of Vienna, Faculty of Computer Science\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Information Science\\\email{bernhard.haslhofer@cornell.edu}}

\begin{document}

\maketitle

\begin{abstract}

The Simple Knowledge Organization System (SKOS) has become the standard model for expressing and publishing controlled vocabularies on the Web. However, SKOS vocabularies are often heterogeneous in terms of quality, which reduces their applicability across system boundaries. Here we investigate how we can support taxonomists in judging the quality of SKOS vocabularies by quality criteria that go beyond the low-level data model integrity constraints defined in the SKOS specification. Based on existing guidelines and best practices for creating controlled vocabularies, we identified existing \textbf{quality criteria}, formalized them into computable \textbf{filtering functions} that identify  affected resources and implemented them in the qSKOS quality assessment tool. We used that tool to analyze existing vocabularies and found that \todo{CM,BH}{summarize results in one sentence}.

\end{abstract}

\input{introduction}

\input{background}

\input{criteria}

\input{analysis}

\input{related_work}

\input{conclusions}

\bibliography{references}
\bibliographystyle{splncs03}

\end{document}