### 2012-03-20 main.tex
minor consistency fixes in secs 1-3

\documentclass{llncs}

\usepackage{url}

\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{booktabs}

\usepackage{color}

\newcommand{\todo}[2]{\textbf{\textcolor{red}{(TODO [#1]: #2)}}}

% metrics = quantitative indicator of sth

\title{Quality Criteria for SKOS Vocabularies}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2}}
\institute{
	University of Vienna, Faculty of Computer Science\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Information Science\\\email{bernhard.haslhofer@cornell.edu}}

\begin{document}

\maketitle

\begin{abstract}

The Simple Knowledge Organization System (SKOS) has become a standard model for controlled Web vocabularies. However, SKOS vocabularies are often heterogeneous in terms of quality, which reduces their applicability across system boundaries. Here we investigate how we can support taxonomists in judging the quality of SKOS vocabularies by quality criteria that go beyond the low-level data model integrity constraints defined in the SKOS specification. We identified existing \textbf{quality criteria}, formalized them into computable \textbf{quality checking functions} that identify affected resources in a given SKOS vocabulary. We implemented these functions in the qSKOS quality assessment tool, analyzed \todo{CM}{XX} existing vocabularies and found possible quality issues in \todo{CM}{YY} vocabularies.

\end{abstract}

\input{introduction}

\input{background}

\input{criteria}

\input{analysis}

\input{related_work}

\input{conclusions}

\bibliography{references}
\bibliographystyle{splncs03}

\end{document}