### 2012-03-22 main.tex
updated abstract

\documentclass{llncs}

\usepackage{url}

\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{booktabs}

\usepackage{color}

\newcommand{\todo}[2]{\textbf{\textcolor{red}{(TODO [#1]: #2)}}}

% metrics = quantitative indicator of sth

\title{Quality Criteria for SKOS Vocabularies}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2}}
\institute{
	University of Vienna, Faculty of Computer Science\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Information Science\\\email{bernhard.haslhofer@cornell.edu}}

\begin{document}

\maketitle

\begin{abstract}

The Simple Knowledge Organization System (SKOS) has become a standard model for controlled vocabularies on the Web. However, SKOS vocabularies often differ in terms of quality, which reduces their applicability across system boundaries. Here we investigate how we can support taxonomists in improving the quality of SKOS vocabularies by pointing out quality issues that go beyond the integrity constraints defined in the SKOS specification. We identified possible \textbf{quality issues} and formalized them into computable \textbf{quality checking functions} that can find affected resources in a given SKOS vocabulary. We implemented these functions in the qSKOS quality assessment tool, analyzed 15 existing vocabularies, and found possible quality issues in all of them.

\end{abstract}

\input{introduction}

\input{background}

\input{criteria}

\input{analysis}

\input{related_work}

\input{conclusions}

\bibliography{references}
\bibliographystyle{splncs03}

\end{document}