### 2012-03-26 conclusions.tex
updated related work and conclusions

%!TEX root = main.tex
\section{Conclusions and Future Work}\label{sec:conclusions}

In this paper we identified possible quality issues in SKOS vocabularies and implemented them as quality checking functions in our qSKOS quality assessment tool. We analyzed a representative set of existing SKOS vocabularies and found  issues in all of them. Labeling and documentation issues were omnipresent and also structural issues, which require further investigation by the vocabulary maintainers, were found in most vocabularies. Although SKOS is designed for Linked Data, many existing vocabularies still resemble their closed-system origin, which results in a relatively low number of in- and out-links. Broken links is a major issue and calls for synchronization mechanisms in order to maintain navigability between concepts.

During our work on qSKOS and while writing this paper we were in contact with some of the creators of the vocabularies we analyzed and also reported back initial results from our analysis. At the time of this writing, we know that our findings lead to improvements in at least two SKOS vocabularies.

We are aware that the quality issues we described in this work are purely quantitative indicators that, on their own, cannot be generalized into statements about the quality of a vocabulary. To learn more about the real-world impact of our work, we would like to conduct a qualitative follow-up study, in which we confront taxonomists with our results. We will also further collect community feedback, enhance the issues list, and set up a Web-based SKOS quality checking service. For larger vocabularies we will also increase the performance and efficiency of our quality checking algorithms.

% \section{Acknowledgements}

% Tom Dent (esd support team) for providing a dump for the IPSV vocabulary
% Andrew Gibson for providing a dump for the Peroxisome knowledge Base and giving valuable feedback on their vocabulary structure.