### 2012-03-27 introduction.tex

%!TEX root = main.tex

\section{Introduction}\label{sec:introduction}

% BACKGROUND

% What is SKOS?
The Simple Knowledge Organization System (SKOS)~\cite{Miles2005} is a standard model for sharing and linking controlled vocabularies on the Web. Many organizations, including the European Union\footnote{EuroVoc, \url{http://eurovoc.europa.eu/}}, the United Nations\footnote{AGROVOC, \url{http://aims.fao.org/website/AGROVOC-Thesaurus/sub}}, or the UK government\footnote{Integrated Public Sector Vocabulary (IPSV), \url{http://doc.esd.org.uk/IPSV}} publish SKOS representations of their controlled vocabularies on the Web so that they can easily be accessed by humans and machines.

% What does "quality" mean in this context?
Each vocabulary reflects domain- and application-specific needs and taxonomists often follow existing standards (cf.,~\cite{ISO25964-1:2011,Z39.19:2005}) or apply their own hand-crafted checks (cf.,~\cite{Coronado2009}) to meet certain quality requirements. However, quality requirements can differ across application domains. The following examples illustrate how quality issues can affect the applicability of SKOS vocabularies for tasks such as query expansion, faceted browsing, or auto-completion:

\begin{itemize}

	\item The SKOS representation of AGROVOC defines concepts in 25 different languages. However, while most concepts have an English description, only 38\% have German labels attached. This can be a problem for multilingual applications that rely on label translations.

	\item An earlier version (v.8.06) of the STW thesaurus contained 5 pairs of concepts with identical labels. As a result, the auto-complete function of the online search interface suggested identical entries, without giving disambiguation information.

% Note: can we replace the LvAK with an example from a public vocabulary? Then we could simply exclude that vocabulary from that paper.

	\item The non-public vocabulary developed by the Austrian Armed Forces (LVAk) contains 11 disconnected concept clusters. When we confronted the thesaurus maintainers with these structures, they recognized them as ``forgotten'' test data that has no practical significance.

    % \item In the DBpedia category\footnote{\url{http://downloads.dbpedia.org/3.7/en/skos_categories_en.nt.bz2}} hierarchy, more than 10 percent of the categories not linked to any other category. When other vocabularies link to the DBPedia categories, e.g., to enable query expansion, it can negatively affect recall and reduce navigability.

\end{itemize}

% PROBLEM / RESEARCH QUESTION

The SKOS specification doesn't make any explicit statements about vocabulary quality. It defines a set of \textbf{integrity conditions} that state ``whether or not given data are consistent with respect to the SKOS data model''~\cite{Miles2005}. Existing quality checking tools, such as the PoolParty Thesaurus Consistency Checker\footnote{\url{http://demo.semantic-web.at:8080/SkosServices/check}} or SKOSify\footnote{\url{http://code.google.com/p/skosify/}} can be used to validate vocabularies against these conditions. However, these conditions fail to capture quality aspects such as those mentioned in the examples before, aspects that go beyond integrity conditions of the SKOS data model. Existing guidelines and manuals consider these aspects but currently rely on \textbf{human judgement}, which is subjective and does not scale for larger, possibly linked vocabularies on the Web.

% CONTRIBUTION

The goal of our work is to help taxonomists in identifying possible quality issues in SKOS vocabularies, which go beyond data model integrity conditions. We want to given them a set of computable quality checking functions that, in combination with the taxonomists' experience and domain expertise, can serve as quality indicators for vocabularies. Finding such quality issues also gives important feedback on the overall vocabulary design process and should, at the end, lead to better vocabularies. The main contributions of this work can be summarized as follows:

\begin{itemize}

	\item We identified 15 \textbf{quality issues} for SKOS vocabularies by examining existing guidelines and formalized them into computable \textbf{quality checking functions} that identify possibly affected resources in a given SKOS vocabulary.

	\item With the \textbf{qSKOS quality assessment tool} we provide a reference implementation of these functions.

	\item We \textbf{analyzed} a representative set of \textbf{15 existing SKOS vocabularies} to learn about possible quality issues in existing Web vocabularies.

\end{itemize}

In the following, we will first discuss what \emph{quality} means in the context of SKOS vocabularies and how it is currently support by the SKOS specification and existing tools. Then we introduce the quality issues we have identified and describe how we implemented them in the qSKOS quality assessment tool. Finally, we report on the results of an analysis we performed on 15 existing SKOS vocabularies and which shows that the quality issues we discussed are real and can lead to the improvement of existing vocabularies.