### 2012-03-28 main.tex

\documentclass{llncs}

\usepackage{url}

\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{booktabs}

\usepackage{array}

\usepackage{color}
%\usepackage{ulem} Why this?

\newcommand{\todo}[2]{\textbf{\textcolor{red}{(TODO [#1]: #2)}}}

%Title ideas:
% - How can we fix quality issues in SKOS vocabularies?
% - Detecting quality issues in SKOS Vocabularies
% - Computable Quality Guidlines for SKOSO Vocabularies
% - ...

\title{Finding Quality Issues in SKOS Vocabularies}
\subtitle{\todo{ALL}{I am still not happy with the title; better ideas?}}
\author{Christian Mader\inst{1} \and Bernhard Haslhofer\inst{2} \and Antoine Isaac\inst{3}}
\institute{
	University of Vienna, Faculty of Computer Science, Austria\\\email{christian.mader@univie.ac.at}
	\and Cornell University, Department of Information Science, USA\\\email{bernhard.haslhofer@cornell.edu}
	\and Vrije Universiteit Amsterdam, The Netherlands\\\email{aisaac@few.vu.nl}
	}

\begin{document}

\maketitle

\begin{abstract}

The Simple Knowledge Organization System (SKOS) has become a standard model for controlled vocabularies on the Web. However, SKOS vocabularies often differ in terms of quality, which reduces their applicability across system boundaries. Here we investigate how we can support taxonomists in improving SKOS vocabularies by pointing out quality issues that go beyond the integrity constraints defined in the SKOS specification. We identified potential \textbf{quality issues} and formalized them into computable \textbf{quality checking functions} that can find affected resources in a given SKOS vocabulary. We implemented these functions in the qSKOS quality assessment tool, analyzed 15 existing vocabularies, and found possible quality issues in all of them.

\end{abstract}

\input{introduction}

\input{background}

\input{criteria}

\input{analysis}

\input{conclusions}

% \section{Acknowledgements}

% Tom Dent (esd support team) for providing a dump for the IPSV vocabulary
% Andrew Gibson for providing a dump for the Peroxisome knowledge Base and giving valuable feedback on their vocabulary structure.

\bibliography{references}
\bibliographystyle{splncs03}

\end{document}