### 2012-03-28 related_work.tex

%!TEX root = main.tex
\section{Related Work}\label{sec:related_work}

% Quality in ontology engineering

% Gangemi~\cite{Gangemi2006} introduced structural, functional and usability-related measures for ontology evaluation, Arpinar~\cite{Arpinar2006} define three types of conflicts that may occur in ontologies, and also Tartir et al.~\cite{Tartir2007} propose a framework for evaluation and ranking of ontologies. Although there are definitely overlaps between ontology evaluation and our research, the quality issues we identified are less tight to underlying data model and formal ontology definition language but, to a large extent, derived from manuals and guidelines taxonomist use in their daily work.