### 2012-03-28 background.tex

%!TEX root = main.tex

\section{Background and Related Work}\label{sec:background}

% Data quality in general
The problem of ``vocabulary quality'' is closely related to the more general one of ``data quality`'' and has been discussed in data and information systems research~\cite{Batini2009}. Pipino et al.~\cite{Pipino2002} argue that dealing with data quality should involve both ``subjective perceptions of the individuals'' and ``objective measurements based on the data''. We see our work as a contribution to the latter.

\paragraph{Quality in Semantic Web and Linked Data Research}

Hogan et. al~\cite{Hogan2010} identify four categories of common errors and shortcomings in RDF documents and Heath and Bizer~\cite{Heath2011} summarize best practices for publishing data on the Web. Ontology evaluation, i.e., measuring the quality of an ontology, has also been discussed extensively~\cite{Vrandecic2010}. However, the authors focus on RDF datasets and ontologies in general. While we could use some criteria suggested here, these need to be completed by considering SKOS-specific properties. \todo{CM, BH}{I suppose that the in-links etc come from that kind of work, no?}

\paragraph{Quality in SKOS Vocabularies}

SKOS specifies in total six \textbf{integrity conditions}~\cite{SkosReference2008}, each of which is a statement that defines under which circumstances data are consistent with the SKOS data model. For example, ``a resource has no more than one value of \texttt{skos:prefLabel} per language tag''. Tools that can check whether these conditions are met are already available: two of the six conditions are defined formally in the OWL representation of SKOS and can therefore be validated by any OWL reasoner. For validating a SKOS vocabulary against the other integrity conditions, one can use tools such as the \textbf{PoolParty Thesaurus Consistency Checker}, or the \textbf{Skosify} validator, which can also correct some detected quality problems.

One issue when assessing the quality of SKOS data is the so-called ``Open World Assumption'', which underlies the Web of Data itself. Established quality notions from closed-world systems, such as referential integrity or schema validation, don't hold anymore, because available information may be incomplete and non-explicitly stated facts cannot be determined as true or false. \todo{AI}{Add ref to DCAM and Pellet, if we have 3 lines available for that.} Work-arounds, often ad-hoc, are thus currently used to evaluate quality in Linked Data sets~\cite{Heath2011}, as is done in the (rule-based) SKOS tools mentioned above.

% MOVE TO RELATED WORK: Gueret et al. [\todo{CM}{find the right reference in google scholar}], for instance, calculate network measures to support human judgement on the quality of Linked Data \todo{CM}{?} graphs.

%In the course of the LATC project, Gueret et al. propose an infrastructure that utilizes crowd-sourcing technology to semi-automatically create links between datasets of the LOD cloud. To ensure the quality of the resulting network, human judgement is supported by calculating a report of locally approximated network measures.

% SKOS Integrity Conditions (http://www.w3.org/TR/skos-reference/):
% S9    skos:ConceptScheme is disjoint with skos:Concept.
% S13   skos:prefLabel, skos:altLabel and skos:hiddenLabel are pairwise disjoint properties. (* NOT expressed formally *)
% S14   A resource has no more than one value of skos:prefLabel per language tag. (* NOT expressed formally *)
% S27   skos:related is disjoint with the property skos:broaderTransitive. (* NOT expressed formally *)
% S37   skos:Collection is disjoint with each of skos:Concept and skos:ConceptScheme.
% S46   skos:exactMatch is disjoint with each of the properties skos:broadMatch and skos:relatedMatch. (* NOT expressed formally *)
% -> 2 of 6 are expressed formally in OWL; the others cannot be expressed because OWL doesn't support disjoint properties

% Poolparty quality checker implementation:
% S13, S14, S27, S46 plus the formally defined S9, S37
% syntactic checks: URI, character encoding, whitespaces
% missing language tags: in SKOS labels and textual content)
% missing labels: prefLabels for skos:Concepts and rdfs:labels for skos:ConceptSchemes)
% loose concepts: concepts that are no top concepts and have no broaders

% SKOSify
% S13, S14, S27

% Mapping qSKOS Integrity Conditions <-> qSKOS quality criteria
% S13 + S14 combined in "Ambigious labels"
% S27 + S46 combined in "Associative vs. Hierarchical Relation Clashes" and "Exact vs. Associative and Hierarchical Mapping Clashes"

\paragraph{Quality in Controlled Vocabularies}

Typical applications of controlled vocabularies are classification, indexing, autocompletion, query reformulation, or serving as a glossary. As we discussed in detail in earlier work~\cite{Nagy2011}, these areas impose specific requirements on vocabulary features, such as structure, availability, and documentation. Quality aspects of controlled vocabularies have already been discussed in standardized guidelines~\cite{ISO25964-1:2011,Z39.19:2005}, manuals~\cite{Svenonius2003,Hedden2010,Aitchison2000,Harpring2010}, tutorials~\cite{Soergel2002}, and scholarly articles \cite{Coronado2009,Soergel1997,Elkin2002,Kless2010}. Our work builds on this literature, but focuses on the less intellectually loaded checks, which can be automatized to assist vocabulary users or publishers. \todo{AI}{In fact Spero's example can be put here}