### 2012-03-30 conclusions.tex
minor modifications in conclusion

%!TEX root = main.tex
\section{Conclusions and Future Work}\label{sec:conclusions}

In this paper we presented possible quality issues in SKOS vocabularies and described how we implemented them as quality checking functions in our qSKOS quality assessment tool. We analyzed a representative set of existing SKOS vocabularies and found  issues in all of them. Labeling and documentation issues were omnipresent and also structural issues, which require further investigation by the vocabulary maintainers, were found in most vocabularies. Although SKOS is designed for Linked Data, many existing vocabularies still resemble their closed-system origin, which results in a relatively low number of in- and out-links. Broken links is a major issue and calls for synchronization mechanisms in order to maintain navigability between concepts in different vocabularies.

During our work on qSKOS and while writing this paper we reported initial results from our analysis to some of the maintainers of the vocabularies we analyzed. At the time of this writing, we know that our findings led to improvements in at least two SKOS vocabularies.

We are aware that the quality issues we described in this work are purely quantitative indicators that, on their own, cannot be generalized into statements about the quality of a vocabulary. To learn more about the real-world impact of our work, we would like to conduct a qualitative follow-up study, in which we confront taxonomists with our results. We will also further collect community feedback, enhance the issues list, and set up a Web-based SKOS quality checking service. For larger vocabularies we also need to increase the performance and efficiency of our quality checking algorithms.