### 2012-05-30 acknowledgements.tex
added ackn and minor comments

%!TEX root = main.tex
\section{Acknowledgements}

The authors would like to thank \todo{CM}{Names} for \todo{CM}{what?}. The work has partly been supported by \todo{CM}{FWF Grant} and an EU Marie Curie International Outgoing Fellowship, grant no. 252206. \todo{CM}{ask Antoine about necessary Ackn.}