### 2012-05-30 acknowledgements.tex
added acknowledgement details

%!TEX root = main.tex
\section{Acknowledgements}

The authors would like to thank Andrew Gibson and Tom Dent for providing dumps of the PXV and IPSV vocabularies as well as Joachim Neubert for valuable feedback on the found issues. The work has partly been supported by the FWF-funded Meketre project (P21571) and an EU Marie Curie International Outgoing Fellowship, grant no. 252206. \todo{CM}{ask Antoine about necessary Ackn.}