### 2012-06-05 acknowledgements.tex
further changes based on antoine's comments

%!TEX root = main.tex
\subsection*{Acknowledgements}

Andrew Gibson and Tom Dent provided PXV and IPSV data and Joachim Neubert, valuable feedback on the found issues. The work is supported by the FWF P21571 Meketre project and EU Marie Curie Fellowship no. 252206.

%The authors would like to thank Andrew Gibson and Tom Dent for providing dumps of the PXV and IPSV vocabularies as well as Joachim Neubert for valuable feedback on the found issues. The work has partly been supported by the FWF-funded Meketre project (P21571) and an EU Marie Curie International Outgoing Fellowship, grant no. 252206.